How to compute the fusion product of MV cycles in type A

### 2021-02-24 draft.tex

    \item The locus $\cW_\mu = G_1\xt[t^{-1}]L_\mu \subset \Grth $.  
    \acom{want $t^\mu\CC[t]$ as opposed to $L_\mu$ here?}
    In modular terms, this corresponds to the locus of those $ (V, \varphi)$ such that $ V $ is isomorphic to the trivial vector bundle on $ \PP$ and such that $ \varphi$ preserves the \ul{Harder--Narasimhan filtration of $V$ at $ \infty$}.  The lattice description of $ \cW_\mu \cap \Grth^+$ is given in Lemma \ref{le:Wmu} below. \acom{que est-ce?} \jcom{It's not very relevant for this paper.  Here is an explanation: every vector bundle on $\PP$ (or maybe any curve) has a special filtration called the HN-filtration; it is a partial flag of subbundles of type determined by $ \mu$.  We require that $ \varphi$ takes this filtration to the standard partial flag at the fibre over $ \infty$.  Maybe ``preserves'' is not the best word above.}

@@

    \item Fix the ``\mu-numeration'' \((e^1_1,\ldots,e^{\mu_1}_1,\ldots,e^1_m,\ldots,e^{\mu_m}_m)\) of the standard basis of $\CC^N$. The Mirkovic--Vybnornov slice $\TT_\mu$ is defined as the set of $A\in M_N(\CC)$ such that 
    \[
        \begin{aligned}
            &\text{for all } 1 \le a,s\le m\,,
            \text{for all } 1\le b\le \mu_a\,, 1\le t\le \mu_s\,, \\
            &\text{if } 1\le t < \mu_s \text{ or } t = \mu_s < b \le \mu_a \\
            &\text{then } (e^t_s)' (A-J_\mu) e^b_a = 0 \,.
        \end{aligned}    
    \]
    In words, $A$ is a $\mu$ Jordan form matrix plus a $\mu\times\mu$ block matrix with possibly nonzero entries occuring in the first $\min(\mu_i,\mu_j)$ columns of the last row of each $\mu_i\times\mu_j$ block. 
    Mirkovic and Vybornov describe $\TT_\mu$ as the slice in which $f$ acts as close to regular nilpotent as possible.
    % $\dim \TT_\mu = N^2 - \dim \OO_\mu$. 
    In pictures, aka by example, the elements $A$ for $\mu=(3,2,2,1)$ look like 
    \[
        \left[\begin{BMAT}(e){ccc;cc;c;c}{ccc;cc;c;c}
             & 1 & & & & & \\
             &  & 1 & & & & \\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{14}^1 \\
              &  & &  & 1 & & \\
              A_{21}^1 & A_{21}^2 & & A_{22}^1 & A_{22}^2 & A_{23}^1 & A_{24}^1 \\
             A_{31}^1 & & & A_{32}^1 & & A_{33}^1 & A_{34}^1 \\
             A_{41}^1 & & & A_{42}^1 & & A_{43}^1 & A_{43}^1
        \end{BMAT}
        \right]
    \]
    Smaller: 
    \[
        \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
            0 & 1 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0\\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2\\
            0 & 0 & 0 & 0 & 1\\
            A_{21}^1 & A_{21}^2 & 0 & A_{22}^1 & A_{22}^2
            \end{BMAT}\right]    
            \mapsto 
            \left[\begin{array}{rr}
                t^{3} - A_{11}^3 t^2 - A_{11}^2 t - A_{11}^1 & -A_{21}^2 - A_{21}^1  \\
                -A_{12}^2 t - A_{12}^1 & t^{2} - A_{22}^2 t - A_{22}^1
            \end{array}\right]
    \]
    % 
    \acom{this ok?} % \jcom{Anne, can you add this definition?}
    \jcom{When you wrote ``In words,'', it is not quite correct because of the diagonal blocks have those 1s.  Also it would be good to have an example.  And what is $f$ in the last sentence? }
    \acom{fixed? As for $f$ I think it is an $\sl_2$ triplet. Hopeful you could explain what MVy meant bc I never quite understood!}

@@

\begin{lemma} \label{le:Grl1l2}
    Let $ L \in \Grth^+ $.  Let $ s \in \AA, s \ne 0 $.  The following are equivalent:
    \begin{enumerate}
        \item $ L $ is in the image of the map $ \Gr^{\lambda_1, \lambda_2}_s \rightarrow \Grth^+$
        \item $ L \in G(\CC[t]) t^{\lambda_1} (t-s)^{\lambda_2}$ \acom{why no $G(\CC[t])$-coset?}
        \item The linear operator $ t $ on $ L_0/L$ has Jordan type $(0,\lambda_1), (s,\lambda_2)$.
    \end{enumerate}
\end{lemma}

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
        \item $ \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$.
    \end{enumerate}
\end{lemma}

Since the above basis will occur often, we will denote it  $ \beta_\mu$.

@@

@@

\jcom{Do we have a reference?}
\acom{Your paper with Sabin??}

>>>>

    \item The locus $\cW_\mu = G_1\xt[t^{-1}]L_\mu \subset \Grth $.  
    \acom{want $t^\mu\CC[t]$ as opposed to $L_\mu$ here?}
    In modular terms, this corresponds to the locus of those $ (V, \varphi)$ such that $ V $ is isomorphic to the trivial vector bundle on $ \PP$ and such that $ \varphi$ preserves the {Harder--Narasimhan filtration of $V$ at $ \infty$}.  The lattice description of $ \cW_\mu \cap \Grth^+$ is given in Lemma \ref{le:Wmu} below.

    \acom{que est-ce?} \jcom{It's not very relevant for this paper.  Here is an explanation: every vector bundle on $\PP$ (or maybe any curve) has a special filtration called the HN-filtration; it is a partial flag of subbundles of type determined by $ \mu$.  We require that $ \varphi$ takes this filtration to the standard partial flag at the fibre over $ \infty$.  Maybe ``preserves'' is not the best word above.}
    % A: Ok thanks!

@@

    \item Fix the ``$\mu$-numeration'' \((e^1_1,\ldots,e^{\mu_1}_1,\ldots,e^1_m,\ldots,e^{\mu_m}_m)\) of the standard basis of $\CC^N$. The Mirkovic--Vybnornov slice $\TT_\mu$ is defined as the set of $A\in M_N(\CC)$ such that 
    \[
        \begin{aligned}
            &\text{for all } 1 \le a,s\le m\,,
            \text{for all } 1\le b\le \mu_a\,, 1\le t\le \mu_s\,, \\
            &\text{if } 1\le t < \mu_s \text{ or } t = \mu_s < b \le \mu_a \\
            &\text{then } (e^t_s)' (A-J_\mu) e^b_a = 0 \,.
        \end{aligned}    
    \]
    In words, $A$ is a $\mu$ Jordan form matrix plus a $\mu\times\mu$ block matrix with possibly nonzero entries occuring in the first $\min(\mu_i,\mu_j)$ columns of the last row of each $\mu_i\times\mu_j$ block. 
    Mirkovic and Vybornov describe $\TT_\mu$ as the slice in which $f$ acts as close to regular nilpotent as possible.
    % $\dim \TT_\mu = N^2 - \dim \OO_\mu$. 
    In pictures, aka by example, the elements $A$ for $\mu=(3,2,2,1)$ look like 
    \[
        \left[\begin{BMAT}(e){ccc;cc;c;c}{ccc;cc;c;c}
             & 1 & & & & & \\
             &  & 1 & & & & \\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2 & A_{13}^1 & A_{14}^1 \\
              &  & &  & 1 & & \\
              A_{21}^1 & A_{21}^2 & & A_{22}^1 & A_{22}^2 & A_{23}^1 & A_{24}^1 \\
             A_{31}^1 & & & A_{32}^1 & & A_{33}^1 & A_{34}^1 \\
             A_{41}^1 & & & A_{42}^1 & & A_{43}^1 & A_{43}^1
        \end{BMAT}
        \right]
    \]
    Smaller: 
    \[
        \left[\begin{BMAT}(e){ccc;cc}{ccc;cc} 
            0 & 1 & 0 & 0 & 0\\
            0 & 0 & 1 & 0 & 0\\
            A_{11}^1 & A_{11}^2 & A_{11}^3 & A_{12}^1 & A_{12}^2\\
            0 & 0 & 0 & 0 & 1\\
            A_{21}^1 & A_{21}^2 & 0 & A_{22}^1 & A_{22}^2
            \end{BMAT}\right]    
            \mapsto 
            \left[\begin{array}{rr}
                t^{3} - A_{11}^3 t^2 - A_{11}^2 t - A_{11}^1 & -A_{21}^2 - A_{21}^1  \\
                -A_{12}^2 t - A_{12}^1 & t^{2} - A_{22}^2 t - A_{22}^1
            \end{array}\right]
    \]
    % 
    \acom{this ok?} % \jcom{Anne, can you add this definition?}
    \jcom{When you wrote ``In words,'', it is not quite correct because of the diagonal blocks have those 1s.  Also it would be good to have an example.  And what is $f$ in the last sentence? }
    \acom{fixed? As for $f$ I think it is an $\sl_2$ triplet. Hopeful you could explain what MVy meant bc I never quite understood!}

@@

\begin{lemma} \label{le:Grl1l2}
    Let $ L \in \Grth^+ $.  Let $ s \in \AA, s \ne 0 $.  The following are equivalent:
    \begin{enumerate}
        \item $ L $ is in the image of the map $ \Gr^{\lambda_1, \lambda_2}_s \rightarrow \Grth^+$
        \item The linear operator $ t $ on $ L_0/L$ has Jordan type $(0,\lambda_1), (s,\lambda_2)$.
        \item $ L \in G(\CC[t]) t^{\lambda_1} (t-s)^{\lambda_2}$ \acom{why no $G(\CC[t])$-coset?}
    \end{enumerate}
\end{lemma}

\begin{proof}
First we recall that for $ L \in \Gr^+$, $ L \in \Gr^{\lambda} = G(\CC[[t]])t^\lambda $ if and only if $ t |_{L_0/L} $ has Jordan type $ \lambda$.

    Now assume that $ L $ lies in the image of the map from $ \Gr^{\lambda_1, \lambda_2}_s$.  We know that $ L(0) \in \Gr^{\lambda_1}, L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on $ \CC[[t]]^m / L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on $ \CC[[t-s]]^m / L(s) $ has Jordan type $ \lambda_2$.  Now, the map $ L_0 / L \rightarrow \CC[[t]]^m / L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \CC[[t]]^m / L(0)$.    This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i).

    On the other hand, if $ L = g t^{\lambda_1} (t-s)^{\lambda_2} $ for some $ g \in G(\CC[t])$, then $ L(0) = g t^{\lambda_1} $ since $ (t-s)^{\lambda_2} \in G(\CC[[t]])$.  In this way, we see that (iii) implies (ii) and the logic can be reversed to get the equivalence.
\end{proof}

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
                \item $ L = Span_\cO(v_1, \dots, v_m)$ for some $ v_i $ of the form $ v_i = t^{\mu_i} e_i + \sum_{j=1} p_{ij} e_j $ where $ p_{ij} \in \CC[t] $ has degree less than $ \min(\mu_i, \mu_j)$.
        \item $ \beta_\mu := \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$.

    \end{enumerate}
\end{lemma}

\begin{proof}
    Let $ L \in \cW_\mu$.  Then $ L = Span_\cO(v_1, \dots, v_m) $ for some $ v_i $ with $ v_i = t^{\mu_i} e_i + \sum_{j=1} q_{ij}t^{\mu_i} e_j $ and $ q_{ij} \in t^{-1} \CC[[t^{-1}]]$.  Since $ L \in \Grth $, we see that $ v_i \in L_0$ which mean that $ p_{ij} := q_{ij}t^{\mu_i} \in \CC[t]$.  If $ \mu_j < \mu_i$, then we can alter this basis to $ v'_i = v_i - r v_j$ for some polynomial $r \in \CC[t]$ and thus ensure that $ p_{ij} $  has degree less than $ \min(\mu_i, \mu_j)$.  Thus (i) implies (ii).

    Suppose that $ L = Span_\cO(v_1, \dots, v_m)$ as in (ii).  Then $[t^{\mu_i} e_i] \in Span_\CC(\beta_\mu) $
in $ L_0/L$.  For this, we can deduce that $ \beta_\mu$ forms a basis for $ L_0/L$.  Hence (ii) implies (iii).  
    Finally, if we know that $ \beta_\mu $ forms a basis, then $ v_i := t^{\mu_i} e_i - \sum_{j=1} p_{ij} e_j \in L $ for some $ p_{ij} \in \CC[t]$ of degree less than $ \mu_j $.  It is easy to see that $ L = Span_\cO(v_1, \dots, v_m) $ and so $ L \in \cW_\mu$.  
    \end{proof}

@@

\begin{proof}
    Let $ L $ lie in this image.  Then $ L = g t^{\mu_1} (t-s)^{\mu_2}$ from some $ g \in N(\CC[t, t^{-1}, (t-s)^{-1}]) $.  Let $ h =(t-s)^{\mu_2} t^{-\mu_2}  $.  Note that $ h \in T_1\CC([[t^{-1}]])$. Then,
    $$ L = h (h^{-1} g h) t^{\mu}$$
    Note that $ h^{-1} g h \in N(\CC((t^{-1})))$ and from KWWY \jcom{Check this}, we know that $ (h^{-1} g h) t^{\mu} \in \cW_\mu$ and so the result follows.
    \end{proof}

@@

\jcom{Do we have a reference?}
\acom{Your paper with Sabin??}
\jcom{I took at a look at Mirkovic-Vybornov (both versions) and at my paper with Sabin.  In Mirkovic-Vybornov, there is nothing about the image of the map.  In the paper with Sabin, we describe the image as $ L \subset L_0$, $[e_1], \dots $ forms a basis and $ z^{\mu_i} e_i \in W_{\mu_i} + L$.  I don't know why I put that third condition.  It seems automatic to me, since from the ``basis'' condition we know that $ W_{\mu_i} + L = L_0$.  Anyway, I think that I will try to write out the proof again for this paper.}
